\documentclass[letterpaper,12pt]{report}
\usepackage[pdftex]{graphicx}
\usepackage{setspace}
\usepackage{fullpage}
\usepackage{pdfpages}
\usepackage[titletoc]{appendix}
\usepackage{titlesec}
\usepackage{paralist}
\usepackage{url}
\usepackage{array}
\usepackage{multirow}
\usepackage{rotating}
\usepackage{xparse}

%% Custom C table column type for fixed width
\newcolumntype{C}[1]{>{\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

% Rotation: \rot[<angle>][<width>]{<stuff>} 
\NewDocumentCommand{\rot}{O{45} O{1em} m}{\makebox[#2][l]{\rotatebox{#1}{#3}}}

%% Adjust title spacing and borders.
\titleformat{\chapter}[display]
{\normalfont\huge\bfseries}{\chaptertitlename\ \thechapter}{20pt}{\Huge}
\titleformat{\section}
{\normalfont\large\bfseries\singlespacing}{\thesection}{1em}{}
\titleformat{\subsection}
{\normalfont\normalsize\bfseries\singlespacing}{\thesubsection}{1em}{}
\titleformat{\subsubsection}
{\normalfont\normalsize\bfseries\singlespacing}{\thesubsubsection}{1em}{}

\titlespacing*{\chapter}{0pt}{5pt}{25pt}
\titlespacing*{\section} {0pt}{3.5ex plus 1ex minus .2ex}{2.3ex plus .2ex}
\titlespacing*{\subsection} {0pt}{3.25ex plus 1ex minus .2ex}{1.5ex plus .2ex}
\titlespacing*{\subsubsection}{0pt}{3.25ex plus 1ex minus .2ex}{1.5ex plus .2ex}

%% Define a new 'leo' style for the package that will use a smaller font.
\makeatletter
\def\url@leostyle{%
  \@ifundefined{selectfont}{\def\UrlFont{\sf}}{\def\UrlFont{\small\ttfamily}}}
\makeatother
%% Now actually use the newly defined style.
\urlstyle{leo}


\title{CLARKSON UNIVERSITY \\
\ \\
Overcoming Geographic Isolation: The Design and Implementation of a Web-based Collaborative Learning Environment \\ 
\ \\
\large{A Thesis By \\
\textbf{Ryan James Lewis} \\
Department of Computer Science \\
}}

\author{Advisor \\
Dr. 	Kathleen Fowler \\
Department of Mathematics \\
\ \\ \\
Submitted in partial fulfillment of the requirements \\
for the degree of \\
Master of Science \\
Computer Science}

\date{December 15, 2011 \\
\vspace{.1in}
Accepted by the Graduate School \\
\vspace{.4in}
\hrulefill \\
Date \hspace{2in} Dean
}


\begin{document}

\maketitle

% Begin signing page
\thispagestyle{empty}
\begin{center}
The undersigned have examined the thesis entitled

\vspace{0.4in}

\textbf{Overcoming Geographic Isolation: The Design and Implementation of a Web-based Collaborative Learning Environment}

\vspace{0.4in}

presented by Ryan James Lewis,

\vspace{0.4in}

a candidate for the degree of \textsc{Master of Science},

\vspace{0.4in}

and hereby certify it is worthy of acceptance.

\vspace{0.5in}
\hrulefill \\

Date \\
\vspace{0.5in}
\hrulefill \\
\hfill Dr. Kathleen Fowler, Advisor

\vspace{0.5in}
\hrulefill \\
\hfill Dr. Christopher Lynch

\vspace{0.5in}
\hrulefill \\
\hfill Dr. David Wick
\end{center}
% End signing page

\begin{abstract}
A web-based Collaborative Environment for IMPETUS is described that is designed to increase the convenience and accessibility of collaboration in Northern New York, an area with a low population density, large land area, and a small amount of federal financial assistance. Students in need of educational assistance, when meeting in-person is not possible, can engage in personalized web-based learning with a tutor using assessments and materials backed by state standards. Educators can create surveys to receive feedback from targets regardless of their location. Parents have access to their children’s data and program information. Additionally, data logged about students and their usage of the system can be used in future analyses and reports. A usability study was performed via a survey distributed to current IMPETUS participants to verify the system design and gather an initial set of qualitative data for future quantification.
\end{abstract}

\cleardoublepage

\pagenumbering{roman}

\begin{small}
\tableofcontents
\listoffigures
\end{small}

\cleardoublepage

\pagenumbering{arabic}

\doublespacing

% Begin chapters
\include{chapter_1/introduction}
\include{chapter_2/software_overview}
\include{chapter_3/software_design}
\include{chapter_4/assessment}
\include{chapter_5/conclusion}
% End chapters

\singlespacing

\raggedright
\bibliography{bibliography/bibliography}
\bibliographystyle{plain}

\doublespacing

\begin{appendices}
%\include{appendix/user_manual}
\include{appendix/analytics}
\end{appendices}

\end{document}
